\documentclass{article}
\setlength\parindent{0pt}
\begin{document}
Kristina Frye\\
CS 350\\
March 11, 2015\\
Individual Project Writeup\\
\\
Laura DeWitt and I worked on the GoodSorts algorithm together. The collaboration worked
out well with both of us sharing the workload to create the final report. As the project
progressed, we ended up settling on six different algorithms to implement: 
Heapsort, Mergesort, Quicksort, Three Way Quicksort, Dual Pivot Quicksort, and Introsort.
We pretty much followed our initial project plan, but our project did evolve as we performed
measurements and make different discoveries about the algorithms we chose.\\

Laura write the code for Quicksort, Three Way Quicksort, and Dual Pivot Quicksort.
I wrote the code for Heapsort, Mergesort, and Introsort. We each wrote descriptions for
each algorithm that were used as the basis for the report as well as examples and pseudocode.\\

I wrote a Java library for performing input/output and averaging functions shared by all of our programs as well as a simple bash script for performing repeated measurements of our programs.
Laura modified that script and tested all the programs against pseudo-random,
sorted, reverse-sorted, almost-sorted, and duplicated data.\\

I tested our programs against the programs downloaded from Princeton as a sanity check to make
sure that our algorithms were correct.\\

After all the measurements were complete, I fleshed out the algorithm descriptions, examples,
and pseudocode, wrote a description for how we performed our testing, and our project
results. Laura did the final proof reading to check for spelling and grammatical errors.
\end{document}